### 2024-12-04 response.tex
add response letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Academic Paper Response Letter. 
% LaTeX Template
% Version 1.0 (20/07/2023)
% 
% This template refers to the toolbox tutorial (in Chinese):
% https://liam.page/2016/07/22/using-the-tcolorbox-package-to-create-a-new-theorem-environment/
% Please refer to the tutorial for your customization. 
% 
% Template license: Creative Commons CC BY 4.0
%
% Author: Yuhan Xie
% Contact email: yhxie0107@gmail.com
% Any revision advice is welcome!
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage{times} % make the section title in Times New Roman

\usepackage{booktabs}
\usepackage{tabu}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=2.5cm,bottom=2cm,left=1.7cm,right=1.7cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{float}
\usepackage{threeparttable} 
\usepackage{multirow}
\usepackage{url}

\usepackage[most]{tcolorbox}
\newcommand{\rrtstar}{$\texttt{RRT}^\texttt{*}$}
\newcommand{\SubItem}[1]{
    {\setlength\itemindent{15pt} \item[-] #1}
}

\newtheorem{constraint}{ADMM constraint}
\newtheorem{example}{Example}

\usepackage[export]{adjustbox} 
\usepackage[capitalise]{cleveref}
\usepackage{xcolor}
\newcommand{\new}[1]{\textcolor{blue}{#1}}
\newcommand{\news}{\color{blue}}
\graphicspath{{Figs/}}

\begin{document}

\newtcbtheorem[auto counter, number within = section]{cmt}{Comment}{
	colbacktitle = black!60!white, colframe = black!60!white,
	colback = black!5!white,
	fonttitle=\bfseries,%fontupper=\itshape,
}{t}

\noindent
Dear Editor and Reviewers:

\vspace{0.2cm}
\noindent
First of all, we would like to express our sincere gratitude to the Editor-in-Chief, Associate
Editor and anonymous reviewers for their valuable comments and suggestions, which have
been incorporated into the revised manuscript such that the revised version is clearer, more
informative, thus more readable. In the revision, all the comments from the reviewers have
been carefully taken into account and thoroughly implemented. Meanwhile, the changes in the
revised version have been highlighted in blue. Our response to the reviewers’ comments (in
1:1 correspondence) is given as follows:
\vspace{0.2cm}
\noindent
Yours sincerely,

%\vspace{0.2cm}
\noindent
Author A, Author B...

\section*{Response to the Editor} 
\begin{cmt*}{}{}
Editor Comments.
\end{cmt*}

\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your response.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Rewiewer 1
\newpage
\section{Response to reviewer 1}

%%%%%%%%%%%%% Comment 1.1
\begin{cmt}{}{}

Comments 1.1

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 1.1

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 1.0
\begin{cmt*}{}{}        %% use * when do not need enumerate

Comments

\end{cmt*}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 1.2
\begin{cmt}{}{}

Comments 1.2

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 1.2

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 1.3
\begin{cmt}{}{}

Comments 1.3

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 1.3

\vspace{0.3cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Rewiewer 2
\newpage
\section{Response to reviewer 2}

%%%%%%%%%%%%% Comment 2.1
\begin{cmt}{}{}

Comments 2.1

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.1

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 2.2
\begin{cmt}{}{}

Comments 2.2

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.2

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 2.3
\begin{cmt}{}{}

Comments 2.3

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.3

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 2.4
\begin{cmt}{}{}

Comments 2.4

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.4

\vspace{0.3cm}

\end{document}