### 2024-12-05 response.tex
update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Academic Paper Response Letter. 
% LaTeX Template
% Version 1.0 (20/07/2023)
% 
% This template refers to the toolbox tutorial (in Chinese):
% https://liam.page/2016/07/22/using-the-tcolorbox-package-to-create-a-new-theorem-environment/
% Please refer to the tutorial for your customization. 
% 
% Template license: Creative Commons CC BY 4.0
%
% Author: Yuhan Xie
% Contact email: yhxie0107@gmail.com
% Any revision advice is welcome!
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
%% Language and font encodings
\usepackage[english]{babel}
\usepackage{times} % make the section title in Times New Roman

\usepackage{booktabs}
\usepackage{tabu}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=2.5cm,bottom=2cm,left=1.7cm,right=1.7cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{float}
\usepackage{threeparttable} 
\usepackage{multirow}
\usepackage{url}

\usepackage[most]{tcolorbox}
\newcommand{\rrtstar}{$\texttt{RRT}^\texttt{*}$}
\newcommand{\SubItem}[1]{
    {\setlength\itemindent{15pt} \item[-] #1}
}

\newtheorem{constraint}{ADMM constraint}
\newtheorem{example}{Example}

\usepackage[export]{adjustbox} 
\usepackage[capitalise]{cleveref}
\usepackage{xcolor}
\newcommand{\new}[1]{\textcolor{blue}{#1}}
\newcommand{\news}{\color{blue}}
\graphicspath{{Figs/}}

\begin{document}

\newtcbtheorem[auto counter, number within = section]{cmt}{Comment}{
	colbacktitle = black!60!white, colframe = black!60!white,
	colback = black!5!white,
	fonttitle=\bfseries,%fontupper=\itshape,
}{t}

\noindent
Dear Editor and Reviewers:

\vspace{0.2cm}
\noindent
First and foremost, we sincerely thank the Editor-in-Chief, Associate Editor, and anonymous reviewers for their insightful comments and suggestions. These have been carefully incorporated into the revised manuscript, resulting in a clearer, more informative, and more readable version. All reviewer comments have been thoroughly addressed, and the corresponding changes have been highlighted in \new{blue} for ease of reference. Our detailed responses to the reviewers' comments are outlined below:
\vspace{0.2cm}
\noindent
Yours sincerely,

%\vspace{0.2cm}
\noindent
Ziqi Yang, Roberto Tron

\section*{Response to the Editor} 
\begin{cmt*}{}{}
Editor Comments.
\end{cmt*}

\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your response.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Rewiewer 1
\newpage
\section{Response to reviewer 1}

%%%%%%%%%%%%% Comment 1.1
\begin{cmt}{}{}

Comments 1.1

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 1.1

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 1.0
\begin{cmt*}{}{}        %% use * when do not need enumerate

Comments

\end{cmt*}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 1.2
\begin{cmt}{}{}

Comments 1.2

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 1.2

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 1.3
\begin{cmt}{}{}

Comments 1.3

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 1.3

\vspace{0.3cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Rewiewer 2
\newpage
\section{Response to reviewer 2}

%%%%%%%%%%%%% Comment 2.1
\begin{cmt}{}{}

Comments 2.1

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.1

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 2.2
\begin{cmt}{}{}

Comments 2.2

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.2

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 2.3
\begin{cmt}{}{}

Comments 2.3

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.3

\vspace{0.3cm}

%%%%%%%%%%%%% Comment 2.4
\begin{cmt}{}{}

Comments 2.4

\end{cmt}
\vspace{0.1cm}
\noindent
\underline{\textbf{Response:}}
\vspace{0.2cm}

Your Response 2.4

\vspace{0.3cm}

\end{document}